\thispagestyle{plain}

\noindent
Automated malaria diagnosis remains a major challenge for researchers. While many deep learning methods exist, most focus on simplified goals like binary classification or just finding infected cells in a blood smear. These specialized models often achieve high scores on benchmarks, but they struggle in real diagnostic settings that require identifying cells, locating parasites, and measuring infection levels all at once.\\

\noindent This thesis proposes a \textbf{Multi-Task Transfer Learning (MTTL)} framework designed to perform a complete diagnostic analysis from single thin blood smear images. The system uses Low-Rank Adaptation (LoRA) and partial fine-tuning to learn four tasks together: multi-class detection, cell segmentation, infection localization, and instance-level classification.

\bigskip

\noindent
Our experiments show that standard Single-Task Learning (STL) works well for simple detection but struggles as complexity increases. We observed the F1 score drop from $0.82$ in simple 1-class tests to $0.38$ in 3-class settings. Our proposed MTTL framework resolves this degradation. When benchmarked against the industry-standard \textbf{YOLOv8 (Nano and Small)} detectors, our model demonstrated superior sensitivity. It achieved an F1-score for infected cells \textbf{32-35\%} higher than the YOLO variants. Furthermore, it reduced parasitemia estimation error to \textbf{1.08\%}, significantly outperforming YOLOv8-Small (3.78\%), YOLOv8-Nano (2.48\%), and our STL baseline (3.30\%), while producing fewer false alarms and sharper localization results.

\bigskip

\noindent
These findings suggest that combining Transfer Learning, Multi-Task Learning, and Parameter-Efficient Fine-Tuning creates a strong foundation for complex medical imaging tasks. This framework improves reliability and makes results easier to interpret, offering a practical strategy for developing diagnostic models that can be adapted to other biomedical fields. The complete implementation and trained models have been released publicly.

\bigskip

\noindent
\textbf{Keywords:} Malaria Diagnosis, Multi-Task Learning, Transfer Learning, Computational Pathology, Deep Learning, Blood Smear Analysis, Object Detection, LoRA, Fine-Tuning

\bigskip

\noindent
\textbf{Code Repository:} \href{https://github.com/GwadeSteve/MTTL-Research-Malaria}{https://github.com/GwadeSteve/MTTL-Research-Malaria}