\thispagestyle{plain}

\noindent 
First, I place my gratitude to God for the strength, health, and guidance granted to me throughout this path.\\

\noindent 
I would like to express my deepest appreciation to my supervisor, \textbf{Prof. Dr. habil. PATRICK NJIONOU SADJANG}. Thank you for your trust, your patience, and the wisdom you shared at every stage of this research. Your mentorship was vital in shaping not just this work, but my growth as an engineer and as a researcher.\\

\noindent 
A special thanks goes to my co-supervisor, \textbf{Dr. ARMIELLE NOULAPEU}, for her sharp eye and honest feedback. Her critiques significantly elevated the quality of this work. I am also indebted to \textbf{LIONEL OWONO}, \textbf{ALEX BRUNO KENFACK}, \textbf{YVAN NGNEUNMEU} and \textbf{Dr. HIOL NICOLAS} for taking the time to review my drafts and offering insights that helped me see things more clearly.\\

\noindent 
To my colleagues and labmates, thank you for the conversations and debates that helped refine the ideas presented here. I also want to give a specific shout-out to my friends \textbf{NYEMB NDJEM EONE ANDRE KEVIN} and \textbf{NJIMEYUP HAROLD FRANCOIS} for their support and feedback.\\

\noindent 
Most importantly, my heart goes out to my family. To my mother, \textbf{NGO HIOLL VERONIQUE}, and my father, \textbf{GWADE ELIAS}, thank you for your endless love and for the sacrifices you made to get me here. Finally, to my girlfriend, thank you for your understanding, and for keeping me motivated during the hardest moments of this journey.